Contracts for unions and intersections are hard for Nickel

### 2021-05-18 paper.tex
Racket highlighting

@@

\newcommand{\nickel}[1]{\lstinline[language=nickel]{#1}}
\newcommand{\racket}[1]{\lstinline[language=lisp]{#1}}

@@

User defined contracts, also called flat contracts, are a particular
kind of contract that behaves similar to an assertion, and whose
main point is to apply some arbitrary predicate (as arbitrary as the
host language allows it) to a value; if the predicate evaluates to
true, the contract passes, otherwise it fails.

For instance, we could define a flat contract \nickel{even} as
\nickel{fun x => x \% 2 == 0}, that checks
whether a number is even by checking the remainder of dividing it by 2.
And then use it to check that a function always returns an
even number, when applied to a \nickel{Number}:

\begin{figure}[h]
\begin{center}
\begin{lstlisting}[language=nickel]
let f = (fun n => n * 2) @ (Number -> even)
\end{lstlisting}
  \end{center}
\caption{A function wrapped with a contract stating that it always returns
an even number}
\label{fig:num-to-even}
\end{figure}

@@

\begin{lstlisting}[language=lisp]
(define/contract overload
(and/c (-> number? number?)
       (-> string? string?))
(lambda (x) x))
\end{lstlisting}
\caption{Non working function overloading on Racket.}

@@

\begin{lstlisting}[language=lisp]
(define/contract united
(or/c (-> number? number?)
      (-> string? string? string?))
(lambda (x) x))
\end{lstlisting}
\caption{Working \racket{or/c} example on Racket.}
\label{code:racket:or/c:working}

@@

\begin{lstlisting}[language=lisp]
(define/contract united
(or/c (-> number? number?)
      (-> even? even?))
(lambda (x) x))
\end{lstlisting}
\caption{Non working \racket{or/c} example on Racket.}
\label{code:racket:or/c:non-working}

@@

\begin{lstlisting}[language=lisp]
(define/contract overcase
  (case-> (-> string? string?)
          (-> number? number? number?)
          )
  (lambda (x [y 0]) (if (number? x)
                   (+ x y)
                   x)))

>>>>

% Copied from https://hal.inria.fr/hal-01140459/file/racket.inc.tex
\lstdefinelanguage{racket} {
  morekeywords=[1]{define, define-syntax, define-macro, lambda, define-stream, stream-lambda},
  morekeywords=[2]{begin, call-with-current-continuation, call/cc,
    call-with-input-file, call-with-output-file, case, cond,
    do, else, for-each, if,
    let*, let, let-syntax, letrec, letrec-syntax,
    define-context, define-controller, Integer, Boolean, get, when-required, when-provided,
    maybe_publish, require, submod, or/c, ->, \#\%module-begin,
    always_publish, with-syntax, define-struct/contract, syntax-case,
    define/contract,
    let-values, let*-values,
    module, provide,
    and, or, not, delay, force,
    \#`, \#',
    \#lang, implement, begin-for-syntax, rename-out,
    quasiquote, quote, unquote, unquote-splicing,
    map, fold, syntax, syntax-rules, eval, environment, query },
  morekeywords=[3]{import, export},
  alsoletter={',`,-,/,>,<,\#,\%},
  morecomment=[l]{;},
%  literate={lambda}{{\lambdaup}}1, % lambda -- look at https://tex.stackexchange.com/questions/119879/math-symbols-in-tt-font
  moredelim=**[is][\color{light-gray}]{<<@<<}{>>@>>},
  moredelim=**[is][\itshape\color{OliveGreen}]{<<;<<}{>>;>>},
  morecomment=[s]{\#|}{|\#},
  sensitive=true,
}

@@

\newcommand{\nickel}[1]{\lstinline[language=nickel]{#1}}
\newcommand{\racket}[1]{\lstinline[language=racket]{#1}}

@@

\unsure{Should we define contracts prior to this?}
User defined contracts, also called flat contracts, are a particular
kind of contract that behaves similar to an assertion, and whose
main point is to apply some arbitrary predicate (as arbitrary as the
host language allows it) to a value; if the predicate evaluates to
true, the contract passes, otherwise it fails.

For instance, we could define a flat contract \nickel{even} as
\nickel{fun x => x \% 2 == 0}, that checks
whether a number is even by checking the remainder of dividing it by 2.
And then use it to check that a function always returns an
even number, when applied to a \nickel{Integer}:

\begin{figure}[h]
\begin{center}
\begin{lstlisting}[language=nickel]
let f = (fun n => n * 2) @ (Integer -> even)
\end{lstlisting}
  \end{center}
\caption{A function wrapped with a contract stating that it always returns
an even number}
\label{fig:num-to-even}
\end{figure}

@@

\begin{lstlisting}[language=racket]
(define/contract overload
(and/c (-> number? number?)
       (-> string? string?))
(lambda (x) x))
\end{lstlisting}
\caption{Non working function overloading on Racket.}

@@

\begin{lstlisting}[language=racket]
(define/contract united
(or/c (-> number? number?)
      (-> string? string? string?))
(lambda (x) x))
\end{lstlisting}
\caption{Working \racket{or/c} example on Racket.}
\label{code:racket:or/c:working}

@@

\begin{lstlisting}[language=racket]
(define/contract united
(or/c (-> number? number?)
      (-> even? even?))
(lambda (x) x))
\end{lstlisting}
\caption{Non working \racket{or/c} example on Racket.}
\label{code:racket:or/c:non-working}

@@

\begin{lstlisting}[language=racket]
(define/contract overcase
  (case-> (-> string? string?)
          (-> number? number? number?)
          )
  (lambda (x [y 0]) (if (number? x)
                   (+ x y)
                   x)))
